\LARGE \bf Enhancing Security in Multi-Robot Systems through Co-Observation Planning on Unsecured Trajectories

### 2023-08-31 paper.tex
finish ADMM section

\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}

\title{\LARGE \bf

@@

\section{Preliminaries}
In this section, we review various mathematical concepts that will provide the foundations and context for our novel constraints.

\subsection{Differentials}
We define the differential of a map $f(x):\real{m}\to\real{n}$ at a point $x_0$ as the unique matrix $\partial_x f \in\real{n\times m}$ such that
\begin{equation}\label{equ:dt_to_dx}
  \left.\dert f\bigl(x(t)\bigr)\right|_{t=0}=\partial_x f\bigl(x(0)\bigr) \dot{x}(0)
\end{equation}
where $t\mapsto x(t)\in\real{n}$ is a smooth parametric curve such that $x(0)=x$ with any arbitrary tangent $\dot{x}(0)$. For later part of this paper, we will use $\dot f$ for $\dert f$ and $\partial_x f$ for $\frac{\partial f}{\partial x}$. The differentials $\partial_x f$ is derived through (\ref{equ:dt_to_dx}) having $\dot f$ divided by $\dot x$.

@@

\subsection{Alternating Directions Method of Multipliers (ADMM)}\label{chapter:ADMM review}
In this section, we briefly review the path planning algorithm based on ADMM constraints discussed in \cite{yangmulti}.
This algorithm can deal with a variety of types of non-convex and non-smooth optimization constraints that are commonly encountered in trajectory optimization tasks once they are formulated in a certain format.
The goal of the section is to lay the groundwork for proposing our novel constraint in the following sections.

Let $x_{ij}\in\real{m}$ denote the position of agent $i\in\{1,\ldots,n\}$ at the discrete-time index $j\in[0,\ldots,T]$, with $m$ being the dimension of the workspace, $n$ the number of agents, and $T$ the planning time horizon. Let $\vx=\stack(\{x_{ij}\})$ denote the aggregated vector of all the agents' trajectories over the entire time interval. The goal of the algorithm is to optimize an objective function $\varPhi(\vx)$ on a set $\Omega$ defined by nonlinear constraints. To solve the problem using ADMM, we introduce a new set of variables $\vz=D(\vx)$ to translate the constraints $\Omega$ on $x$ to (typically lower-dimensional) sets of constraints on $\vz$, such that the projection operator on $\zeta$ can be easily implemented (e.g., a constrained distance between two agents is translated into their difference constrained to an origin-centered sphere).
The new set of constraints $\zeta$ on $z$ is incorporated using an indicator function $\varTheta_\zeta$. The main problem is written as:
\begin{equation}\label{eq:ADMMSetConstraint_general}
		\begin{array}{r@{\quad}c}
    	\min&\varPhi(\vx)+\varTheta_\zeta(\vz) \\
        s.t.& D(\vx)-\vz=0,
      \end{array}
  \end{equation}
  where $D(\vx)= [D_1(\vx)^T,\dots,D_l(\vx)^T]^T$ is a vertical concatenation of different functions for different constraints.

  And having the update steps shown as:
  \begin{subequations}\label{eq:ADMMupdate}
    \begin{align}
      \label{eq:admm-mod-update-q}\vx^{k+1}&:=\argmin_\vx(\varPhi(\vx^k) +\frac{\rho}{2}\norm{D(\vx)-\vz^k+\vu^k}_2^2)\\
      \vz^{k+1}&:=\Pi_\zeta(D(\vx^{k+1})+\vu^k) \label{eq:ADMM_constraint_update}\\
      \vu^{k+1}&:=\vu^k+D(\vx^{k+1})-\vz^{k+1},
    \end{align}
  \end{subequations}

  Since $D(\vx)$ is the vertical concatenation of $D_l(\vx)$, the projection of each set $\zeta_l$ is independent for each constraint and can be computed separately. The advantage of this formulation is that we can choose $D(\vx)$ such that the new constraint set~$\zeta$ becomes simple to compute; however, the drawback is that we move the non-convexity to the primal cost function, i.e., in the update for $\vx$ in \eqref{eq:admm-mod-update-q}.
Additionally, the function~$D(\vx)$ can be used to select only the subset of the variables on which a constraint depends.

We consider the following types of traditional path planning constraints:
  \subsubsection{Velocity constraints}
  The movement of each agent is constrained by a maximum distance in any direction over a single discrete time step. We then define the function $D(\vx)$ to return the velocity vectors for the $i$-th agent at time step $j$:
\begin{equation}
     D_{ij}(\vx) = x_{ij}-x_{i(j-1)}, \quad j\in\{1,\ldots, T\}
\end{equation}
and the constraint set is:
\begin{equation}\label{eq:velocity_constraint}
   \zeta_{ij} =\bigl\{z \mid \norm{z} \leq v_{max}\bigr\}.
\end{equation}
The projection operator $\Pi_\zeta(z)$ for this constraint implies that the projection of the vector $z$ is inside a sphere with a radius of $v_{max}$ which can then be written as:
\begin{equation}\label{eq:velocity_constraint_update}
    \Pi_{\zeta}(z) = \begin{cases}
 v_{max}\frac{z}{\norm{z}} &\text{if } \norm{z} > v_{max},\\
z	& otherwise.
\end{cases}
\end{equation}

This constraint could be adapted also to enforce more refined dynamical models.

  \subsubsection{Convex obstacles}: We use convex polygons to model regions defining the boundary of the workspace, as well as solid obstacles that cannot be entered by agents (note that the latter represent a non-convex constraint set from the point of view of the optimization problem). Obstacles with non-convex shapes can be modeled using unions of the convex sets.

  We enforce the constraints at each discrete time step $x_{ij}$ (enforcement of the constraints between these points can be achieved by making the obstacles sightly bigger than its actual size). For a convex area, waypoints stays unchanged it they are outside the zone, and if the waypoints are inside the zone, the constraint function $D(\vx)$ returns the least negative distance from the waypoints to the zone defined by hyperplanes which is represented as:
\begin{equation}\label{eq:zoneConstraintMap}
D_i(\vx) = \bmat{d_{11} \\\vdots\\ d_{nm}}
\end{equation}
where 
\begin{align}
    d_{ij}&= \max(\min(d_{ijk},0)),\\
   d_{ijk} &= p_{ij}\transpose n_k - m_k,
\end{align}
and $n_k$ is the normal vector and $m_k$ is the scalar offset defining the $k$-th hyperplane.
The corresponding constraint set is simply
\begin{equation}
    \zeta= \{z \mid z = 0 \},
\end{equation}
with a projection function:
\begin{equation}\label{eq:zoneConstraintProjection}
    \Pi_\zeta(z) = 0.
\end{equation}
The motivating idea for \eqref{eq:zoneConstraintMap} and \eqref{eq:zoneConstraintProjection} is to find waypoints inside the zone and project them to the closest boundary. As mentioned before, non-convex obstacles can be handled by the union of (possibly overlapping) convex obstacles.
\subsubsection{Waypoints with flexible deadlines}
In this type of constraint, we assume that an agent needs to go through a spherical ball around a given point $p$ with a radius of $d_{max}$ at some time instant $j$ belonging to a given time window $[t_1,t_2]$. In this case, we define the function $D(\vq)$ to return the smallest distance from the point $p$ to any point on the trajectory restricted to the relevant time window. This can be expressed precisely in the following form:
\begin{equation}\label{eq:waypoint_constraint_D}
D_i(\vq) = \min_{i\in\{1\dots,n\}, j\in\{t_1,\dots,t_2-1\}}\bigl(\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})\bigr),
\end{equation}
with the constraint set:
\begin{equation}\label{eq:waypoint_constraint}
\zeta= \{z \mid z < z_{max} \},
\end{equation}
where $\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})$ returns the distance between the fixed point $p$ and the segment $\overrightarrow{q_{ij}q_{i(j+1)}}$. Note that this function returns the smallest distance between $(p,q_{ij})$ and $(p,q_{i(j+1)})$ if the projection of the point $p$ does not lie on the line segment~$\overrightarrow{q_{ij}q_{i(j+1)}}$; as a consequence, this constraint does not need to be satisfied exactly at one of the points on the discretized trajectory, but it can also be satisfied ``en route'' on the segment between them.
The projection $\Pi_\zeta(z)$ for this constraint can be written as:
\begin{equation}\label{eq:waypoint_constraint_update}
\Pi_\zeta(z) = \min(z,z_{max}).
\end{equation}
Note that \eqref{eq:waypoint_constraint} and \eqref{eq:waypoint_constraint_update} are equivalent to \eqref{eq:velocity_constraint} and \eqref{eq:velocity_constraint_update}, except for the fact that the quantity in \eqref{eq:waypoint_constraint_D} is always a positive scalar.

  \subsection{Householder rotations}\label{sec:householder}

  To formulate the operator $\Pi_\zeta$ for our novel constraint, we will define a differentiable transformation of the constraint set to a canonical form. This transformation will include a rotation that we derive from a modified version of Householder transformations \cite{householder1958unitary}. With respect to the standard definition, our modification ensures that the final operator is a proper rotation (i.e., not a reflection). We call our version of the operator a \emph{Householder rotation}. In this section we derive Householder rotations and their differentials for the 3-D case; the 2-D case can be easily obtained by embedding it in the $z=0$ plane.

@@

  We compute the differential of $H$ implicitly using the relation \eqref{equ:dt_to_dx}. We will use the notation $\cross{v}:\real{3}\to\real{3\times 3}$ to denote the matrix representation of the cross product with the vector $v$, i.e.,
  \begin{equation}
    \bmat{v_1\\v_2\\v_3}\mapsto \bmat{0 & -v_3 &v_2\\v_3 & 0 & -v_1\\-v_2 & v_1 & 0},
  \end{equation}
  such that $[v]_\times w=v\times w$ for any $w\in\real{3}$. One can verify by direct computation the following property:
  \begin{equation}\label{eq:asymmetric to cross}
    wv\transpose-vw\transpose=\cross{\cross{v}w}.
  \end{equation}
  \begin{proposition}\label{prop:Hderivitive}
    Let $\nu_\cF(t)$ represent a parametric curve. Then we have
    \begin{equation}
      \dot{H}=H\cross{-2M\dot{\nu}_{\cF}}
    \end{equation}
    where the matrix $M\in\real{3\times3}$ is given by
    \begin{equation}
      M=[u]_{\times}  \frac{ \left( I - u u\transpose \right) \left( I- \nu_\cF \nu_\cF\transpose \right)} {\norm{u'} \norm{\nu_\cF}}.
    \end{equation}
  \end{proposition}

\subsection{Rapidly-exploring Random Trees}
We use RRT* as a component of the solution \cite{karaman2010incremental}, which is an optimized variant of the rapidly-exploring Random Trees (RRT). As a optimal path planning algorithm, RRT* returns the shortest paths between an initial location and multiple potential goals. We assume that the generated paths can be travelled in both directions (this is used later in our analysis). The basic RRT* algorithm is shown in Algorithm \ref{alg:rrt}. For a given tree $G=(V,E)$, key functions called by the algorithm are
\begin{description}
\item[$\texttt{Sample}(i)$] This function returns independent identically distributed samples from the free configuration space.
\item[$\texttt{Nearest}(x)$] This function returns the vertex $v\in V$ that is closest in Euclidean distance to the input point $x$.
\item[$\texttt{Steer}(x,y)$] This function returns a point $z$ that minimize $\norm{z-y}$ subject to $\norm{z-x}\leq \eta$ for a prespecified $\eta>0$.
\item[$\texttt{ObstacleFree}(x,y)$] This function returns whether the line segment between $x$ and $y$ is free of collision.
\item[$\texttt{Cost}(v)$] This function assigns a non-negative cost (total travel distance in our application) to the unique path from the initial position to $v$ . 
\item[$\texttt{Parent}(v)$] This is a function that maps the vertex $v$ to $v'\in V$ such that $(v',v)\in E$
\end{description}

\section{Secure Multi-robot path planning using co-observation and reachability constraints}

Besides traditional path planning constraints, additional security constraints need to be incorporated into the path planning problem to prevent cases where an attacker changes a robot's path between two observations to reach a forbidden region without being detected. The work of \cite{wardega2019resilience} provide exact solution for grid-world configuration. Using the onboard sensing capabilities, robots can perform inter-robot observations as an additional security measure. Each robot' self-report includes not only its own status, but also additional observations of other robots' status. For example, robot $i$'s self-report may include its observation of robot $j$, ``$i$ report its arrival at $v$ and $i$ observes $j$ at location $w$''. Similarly, robot $j$ also reports its observations of robot $i$ to the CE. This solution provides paths with an observation schedule such that any attempts by a compromised robot to violate the safety constraints would necessarily break the observation plan and be detected.

We formulate a way to enforce an empty intersection between forbidden regions and ellipsoidal reachability region while optimizing the trajectory, such that if an attacker takes control of the robots, they cannot perform an undetected attack without breaking the co-observation schedule. To the best of our knowledge, our path planning algorithm is the only one flexible enough to handle the safety related spatio-temporal constraints introduced by co-observations and reachability.

\subsection{Co-observation schedule}
The co-observation constraints is used to guarantee that, at time required by the schedule, two robots should get close enough with each other to observe each other's behavior. The co-observation constraint is modeled as a relative distance constraint between two robots at some time instant $j$ (i.e., to require two agents see each other within a certain radius to inspect each other, or to exchange data). The location of the co-observation location should make sure that the ellipsoidal reachability region between each co-observation, considered in later sections \ref{sec:reachability}, has an empty intersection with all forbidden regions.
 We write the function for the constraint as:
\begin{align}
D(\vq) = \overrightarrow{q_{aj}q_{bj}},
\end{align}
where $a,b$ are the indices of the pair of agents required for a mutual inspection. The corresponding constraint set is simply
\begin{equation}
\zeta= \{z \mid z \leq d_{max} \},
\end{equation}
with a projection function:
\begin{equation}\label{eq:co-observation_constraint_update}
\Pi_{\zeta}(z) = \begin{cases}
d_{max}\frac{z}{\norm{z}} &\text{if } \norm{z} > d_{max},\\
z	& otherwise.
\end{cases}
\end{equation}
The locations $q_{aj}$ and $q_{bj}$ where the co-observation performed are computed as part of the optimization.

\subsection{Reachability constraints}\label{sec:reachability}
  In this section, we first provide the definition of an \emph{ellipsoidal reachability region}. We then provide a differentiable map to transform such region in a canonical axis-aligned form where the operator $\Pi_\zeta$ and its differential can be obtained; these operators are then extended to the general case via the aforementioned transform. The overall goal is to define the functions $D(q)$, its differential, and the operator $\Pi_\zeta$ for \emph{ellipsoidal reachability regions} with respect to the forbidden regions that can then be used in the ADMM formulation in Chapter \ref{chapter:Multi-agent-ADMM}.

\subsection{Definition of reachability constraint}\label{sec:ellipsoidal definition}
The reachability region is defined as the set of locations $q(t)$ that a robot can reach between two given fixed positions:

\begin{definition}
	The \emph{reachability region} for two waypoints $q(t_1)=q_1$, $q(t_2)=q_2$ is defined as the sets of points $q'$ in the workspace such that there exist a trajectory $q(t)$ where $q(t')=q'$, $t_1\leq t' \leq t_2$ and $q(t)$ satisfies the velocity constraint $d(q(t),q(t+1))\leq v_{max}$.
\end{definition}
This region can be analytically bounded via an ellipsoid:
\begin{definition}\label{def:Reachability}
	The \emph{reachability ellipsoid} is defined as the region  $\mathcal{E}(q_1,q_2,t_{1},t_{2})=\{\tilde{q}\in\mathbb{R}^n: d(q_1,\tilde{q})+d(\tilde{q}+q_2)<2a\}$, where $a=\frac{v_{maq}}{2}(t_2-t_1)$.
\end{definition}

@@

The reachability ellipsoid is an over-approximation of the exact reachability region; the difference between the two is due to the discretization of the trajectory, and the fact that $\cE$ does not consider the presence of obstacles.
\subsection{Transformation to canonical coordinates}\label{sec:rotation2Standard}
  To simplify the problem, a canonical rigid body transformation is used to transform the ellipse $\cE$ from a global frame $\cF$ to a canonical frame $\cF_\cE$. The latter is defined such that the center of the ellipsoid is located at the origin and the foci are aligned with the first axis of $\cF_\cE$. Since the transformation depends on the two foci, i.e., two waypoints of an agent, the challenge here is to construct the transformation in a differentiable way.
  For the convenience of derivation, we define the coordinate transformation from $\cF$ to $\cF_\cE$ using a rotation $R^\cF_\cE$ and a translation $o^\cF_\cE$, which, to simplify the notation, from now on we simply refer to as $R$ and $o$, respectively. The transformation of a point from the frame $\cF_\cE$ to the frame $\cF$ and its inverse are given by the formulation:
  \begin{equation}
    \Fframe{q}=R\Eframe{q}+o,\label{eq:E to F transform}\quad
    \Eframe{q}=R\transpose(\Fframe{q}-o).
  \end{equation}
  The reverse transformation is $\Eframe{q}=R\transpose( \Fframe{q}-o)$.

>>>>

\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\def\sZ{\mathcal{Z}}
\title{\LARGE \bf

@@

\section{Multi-robot trajectory planning with co-observation}

In this section, we aims to enhance the security of MRSs against malicious takeovers and deviations by attackers. Specifically, we design an inter-robot observation plan (co-observation schedule) to ensures that, during the task period, robots are constantly in proximity to one another according to a schedule, in order to observe and detect potential hazardous behaviors. This achieved by keeping the potential region that the robots can possibly reach between each consecutive co-observations away from the forbidden regions. In this way, the plan ensure that any potential deviations to these forbidden regions will cause the corresponding robot to miss their next co-observation with other robots. \cite{wardega2019resilience} solved this problem in grid world for robots with fixed start and goal locations (Figure~\ref{fig:Grid-example-application}) but was unable to optimizing over arbitrary smooth cost functions.

We formulate the planning problem into an optimal trajectory optimization problem to minimize arbitrary smooth objective functions. The security constraints including co-observation schedule and reachability analysis, alongside with traditional path planning constraints, are transformed into different types of spatio-temporal constraints in the optimization problem. We denote as $q_{ij}\in\real{m}$ the position of agent $i$ at the discrete-time index $j$, with $m$ representing the dimension of the state space. For a team of $n$ agents, and a task time horizon of $T$, the overall trajectory of the multi-agent system can be represented as an aggregated vector $\vq\in \real{n m T}$. 
The goal of our path planning problem is to minimize or maximize an objective function $\varPhi(\vq)$ under a set of nonlinear constraints described by a set $\Omega$, formally:
\begin{equation}\label{eq:general-problem}
	\begin{split}
		\min/\max & \quad \varPhi(\vq)\\
		\textrm{subject to} &\quad \vq \in \Omega.
	\end{split}
\end{equation}

As an example application, we assume that there exists an underlying slowly-time-varying scalar or vector field $\vx$ whose value needs to be estimated at a given number of discrete locations. To explore unknown spatially-distributed fields, autonomous agents are usually required to traverse the entire unknown environment while collecting sensory data to estimate a corresponding map (Figure~\ref{fig:ADMM-example-application}). We model the map as a set of locations of interests arranged on a regular grid, where at each location, we assume that there is a corresponding slowly-time-varying quantity. For our purposes, we are interested in finding paths that best reconstruct the field, i.e., that achieve the minimum uncertainty. We define the cost function $\varPhi(\vq)$ to be an approximation of the uncertainty at any of the given points, and we aim to minimize such maximum uncertainty. We focus on a particular choice of objective function $\varPhi(\vq)$, in the context of a mapping and estimation applications, for the optimization problem \eqref{eq:general-problem}. \zyang{Do we need detail of the problem formulations?}

\begin{figure}
     \centering
     \subfloat[Continuous world trajectory \label{fig:ADMM-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 2cm 2cm 2cm]{Example_application}}
     \subfloat[Grid-world trajectory. \label{fig:Grid-example-application}]{\includegraphics[width=0.45\linewidth, trim = 2cm 0.5cm 0cm 2cm]{Kacper_result}}
     \caption{Trajectory design of an example map exploration task for a three-robot system with start locations and destinations fixed. Task are planned in a $8 \times 8$ grid world. Zone 1 is obstacle, Zone 2 and Zone 3 are safe zones. Fig~\ref{fig:Grid-example-application} shows the planned trajectory with co-observation schedule in grid-world. Fig~\ref{fig:ADMM-example-application} shows the trajectory design in continuous world optimized with respect to a map exploration task}
     \label{fig:example-application}
 \end{figure}

 This section is organized as follows. We begin by presenting the foundational mathematical concepts, notably the basic ADMM optimal trajectory solver. Subsequently, we seamlessly integrate security constraints, encompassing co-observation schedule and reachability constraints, into our proposed solver. Concluding this section, we present the obtained results alongside identifying limitations that will be addressed in the upcoming section.

\subsection{Preliminary}
In this section, we review various mathematical concepts that will provide the foundations and context for the solver.

\subsubsection{Differentials}
We define the differential of a map $f(x):\real{m}\to\real{n}$ at a point $x_0$ as the unique matrix $\partial_x f \in\real{n\times m}$ such that
\begin{equation}\label{equ:dt_to_dx}
  \left.\dert f\bigl(x(t)\bigr)\right|_{t=0}=\partial_x f\bigl(x(0)\bigr) \dot{x}(0)
\end{equation}
where $t\mapsto x(t)\in\real{n}$ is a smooth parametric curve such that $x(0)=x$ with any arbitrary tangent $\dot{x}(0)$. For later part of this paper, we will use $\dot f$ for $\dert f$ and $\partial_x f$ for $\frac{\partial f}{\partial x}$. The differentials $\partial_x f$ is derived through (\ref{equ:dt_to_dx}) having $\dot f$ divided by $\dot x$.

@@

\subsubsection{Alternating Directions Method of Multipliers (ADMM)}\label{chapter:ADMM review}
The basic idea behind ADMM based solver \cite{zyang} is to split the constraints from the objective function using a different set of variables $\vz$, and then solve an Augmented Lagrangian formulation of the optimization problem in \eqref{eq:general-problem}.
More specifically, we can rewrite the constraint $\vq\in\varOmega$ using an indicator function $\varTheta$, and include it in the objective function. However, in the traditional application of ADMM, the duplicated variables $\vz$ are solved through a projection to the constraint set $\Omega$. However, in path planning problems, some constraints are non-convex, rendering the projection step more difficult (due to the presence of multiple local minima).
To deal with this problem, we propose a minor generalization of the ADMM formulation~\eqref{equ:ADMMSetConstraint} where we allow $\vz$ to replicate an arbitrary function of the main variables $\vq$ (instead of being an exact copy): 
\begin{equation}\label{eq:ADMMSetConstraint_modified}
	\begin{aligned}
		&\max\quad \varPhi(\vq)+\varTheta(\vz) \\
		& \begin{array}{r@{\quad}c}
			s.t.& D(\vq)-\vz=0
		\end{array} 
	\end{aligned}
\end{equation}
where $D(\vq)= [D_1(\vq)^T,\dots,D_l(\vq)^T]^T$ is a vertical concatenation of different functions for different constraints.

The update steps of the ADMM algorithm can be then denoted as \cite{Boyd2011}:
\begin{subequations}\label{eq:ADMMupdate}
	\begin{align}
		\label{eq:admm-mod-update-q}\vq^{k+1}&:=\argmin_\vq(\varPhi(\vq^k) +\frac{\rho}{2}\norm{D(\vq)-\vz^k+\vu^k}_2^2)\\
		\vz^{k+1}&:=\Pi_\mathcal{Z}(D(\vq^{k+1})+\vu^k) \\
		\vu^{k+1}&:=\vu^k+D(\vq^{k+1})-\vz^{k+1},
	\end{align}
\end{subequations}
where $\Pi_\mathcal{Z}$ is the new projection to the modified constraint set $\mathcal{Z}$, $\vu$ represents a scaled dual variable that intuitively accumulates the sum of primal residuals
\begin{equation}\label{eq:primal-residual}
	\vr^{k}=D(\vq^{k+1})-\vz^{k+1}.
\end{equation}
Checking the primal residuals alongside with the dual residuals 
\begin{equation}\label{eq:dual-residual}
	\vs^{k}=-\rho(z^{k}-z^{k-1}),
\end{equation}
after each iteration, the steps are reiterated until convergence when the primal and dual residuals are small, or divergence when primal and dual residual remains large after a fixed large number of iterations.
Since $D(\vq)$ is the vertical concatenation of $D_i(\vq)$, the projection of each set $\mathcal{Z}_i$ is independent for each constraint and can be computed separately. The advantage of this formulation is that we can choose $D(\vq)$ such that the new constraint set~$Z$ becomes simple to compute; however, the drawback is that we move the non-convexity to the primal cost function, i.e., in the update for $\vq$ in \eqref{eq:admm-mod-update-q}.
Additionally, the function~$D(\vq)$ can be used to select only the subset of the variables on which a constraint depends.

We consider the following types of traditional path planning constraints in the ADMM constraints:
\subsubsection*{Velocity constraints}
  The movement of each agent is constrained by a maximum distance in any direction over a single discrete time step. We then define the function $D(\vq)$ to return the velocity vectors for the $i$-th agent at time step $j$:
\begin{equation}
     D_{ij}(\vq) = q_{ij}-q_{i(j-1)}, \quad j\in\{1,\ldots, T\}
\end{equation}
and the constraint set is:
\begin{equation}\label{eq:velocity_constraint}
   \sZ_{ij} =\bigl\{z \mid \norm{z} \leq v_{max}\bigr\}.
\end{equation}
The projection operator $\Pi_\zeta(z)$ for this constraint  is simply a projection of $z$ to the interior of a sphere centered at the origin and having radius $v_{max}$. This constraint could be adapted to enforce more refined dynamical models.

%implies that the projection of the vector $z$ is inside a sphere with a radius of $v_{max}$ which can then be written as:
%\begin{equation}\label{eq:velocity_constraint_update}
%    \Pi_{\zeta}(z) = \begin{cases}
% v_{max}\frac{z}{\norm{z}} &\text{if } \norm{z} > v_{max},\\
%z	& otherwise.
%\end{cases}
%\end{equation}

\subsubsection*{Convex obstacles}

We use convex polygons to model regions defining the boundary of the workspace, as well as solid obstacles that cannot be entered by agents (note that the latter represent a non-convex constraint set from the point of view of the optimization problem). Obstacles with non-convex shapes can be modeled using unions of the convex sets.

The convex region $\cP$ is defined using a collection of $l$ hyperplanes $n_{k} q = m_{k}$, where $n_k$ is the normal vector (outward pointing for obstacle boundaries and inward pointing for workspace boundaries) and $m_k$ is the scalar offset defining the $k$-th hyperplane. The distance between a single waypoint $q_{ij}$ and the $k$th hyperplane $d_{k}(q_{ij},\cP)$ is defined as:
\begin{equation}
    d_{k}(q_{ij},\cP)= p_{ij}\transpose n_k - m_k.
\end{equation}

The constraint function $D(q_{ij})$ returns the least negative distance from $q_{ij}$ to all hyperplane boundaries of region $\cP$, defined as:
\begin{equation}\label{eq:zoneConstraintMap}
	D(q_{ij}) = \max_{k=1,\dots,l} (\min(d_{k}(q_{ij},\cP),0))
\end{equation}

%\rtron{Define for a single $x_{ij}$, then just say that we add constraints for all $ij$. Instead of using $min(d_{ijk})$, use $dist(x_{ij},\cP)$, where $\cP$ is the cvx polygon} We enforce the constraints for all $ij$ (enforcement of the constraints between these points can be achieved by making the obstacles sightly bigger than its actual size).

The corresponding constraint set is simply
\begin{equation}
    \sZ= \{z \mid z = 0 \},
\end{equation}
with a projection function:
\begin{equation}\label{eq:zoneConstraintProjection}
    \Pi_\sZ(z) = 0.
\end{equation}
The motivating idea for \eqref{eq:zoneConstraintMap} and \eqref{eq:zoneConstraintProjection} is to find waypoints inside the zone and project them to the closest boundary. As mentioned before, non-convex obstacles can be handled by the union of (possibly overlapping) convex obstacles.

\subsubsection*{Waypoints with flexible deadlines}
We assume that an robot $i$ needs to reach a given point $p$ within a radius of $d_{max}$ at some time instant $j$ belonging to a given time window $[t_1,t_2]$. The constraint function $D(\vq)$ to return the smallest distance from the point $p$ to waypoint on the trajectory restricted to the relevant time window. This can be expressed precisely in the following form:
\begin{equation}\label{eq:waypoint_constraint_D}
D(\vq) = \min_{j\in\{t_1,\dots,t_2-1\}}\bigl(\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})\bigr),
\end{equation}
with the constraint set:
\begin{equation}\label{eq:waypoint_constraint}
\sZ= \{z \mid z < z_{max} \},
\end{equation}
where $\dist(p,\overrightarrow{q_{ij}q_{i(j+1)}})$ returns the distance between the fixed point $p$ and the segment $\overrightarrow{q_{ij}q_{i(j+1)}}$. Note that this function returns the smallest distance between $(p,q_{ij})$ and $(p,q_{i(j+1)})$ if the projection of the point $p$ does not lie on the line segment~$\overrightarrow{q_{ij}q_{i(j+1)}}$; as a consequence, this constraint does not need to be satisfied exactly at one of the points on the discretized trajectory, but it can also be satisfied ``en route'' on the segment between them.
The projection $\Pi_\sZ(z)$ for this constraint can be written as:
\begin{equation}\label{eq:waypoint_constraint_update}
\Pi_\sZ(z) = \min(z,z_{max}).
\end{equation}
%Note that \eqref{eq:waypoint_constraint} and \eqref{eq:waypoint_constraint_update} are equivalent to \eqref{eq:velocity_constraint} and \eqref{eq:velocity_constraint_update}, except for the fact that the quantity in \eqref{eq:waypoint_constraint_D} is always a positive scalar.

\rtron{Give example application here, and include application of ADMM without the security constraints, but highlighting a path that would break the security}
\zyang{plot new simpler example, add collision avoidance}

 \subsubsection{Householder rotations}\label{sec:householder}

  To formulate the operator $\Pi_\sZ$ for our novel constraint, we will define a differentiable transformation of the constraint set to a canonical form. This transformation will include a rotation that we derive from a modified version of Householder transformations \cite{householder1958unitary}. With respect to the standard definition, our modification ensures that the final operator is a proper rotation (i.e., not a reflection). We call our version of the operator a \emph{Householder rotation}. In this section we derive Householder rotations and their differentials for the 3-D case; the 2-D case can be easily obtained by embedding it in the $z=0$ plane.

@@

  We compute the differential of $H$ implicitly using the relation \eqref{equ:dt_to_dx}. We will use the notation $\cross{v}:\real{3}\to\real{3\times 3}$ to denote the matrix representation of the cross product with the vector $v$, i.e.,
  \begin{equation}
    \bmat{v_1\\v_2\\v_3}\mapsto \bmat{0 & -v_3 &v_2\\v_3 & 0 & -v_1\\-v_2 & v_1 & 0},
  \end{equation}
  such that $[v]_\times w=v\times w$ for any $w\in\real{3}$. %One can verify by direct computation the following property:
  %\begin{equation}\label{eq:asymmetric to cross}
  %  wv\transpose-vw\transpose=\cross{\cross{v}w}.
  %\end{equation}
  \begin{proposition}\label{prop:Hderivitive}
    Let $\nu_\cF(t)$ represent a parametric curve. Then we have
    \begin{equation}
      \dot{H}=H\cross{-2M\dot{\nu}_{\cF}}
    \end{equation}
    where the matrix $M\in\real{3\times3}$ is given by
    \begin{equation}
      M=[u]_{\times}  \frac{ \left( I - u u\transpose \right) \left( I- \nu_\cF \nu_\cF\transpose \right)} {\norm{u'} \norm{\nu_\cF}}.
    \end{equation}
  \end{proposition}

%\subsection{Secure Multi-robot path planning using co-observation and reachability constraints}
\subsection{Co-observation schedule constraint}
The co-observation constraints is used to guarantee that, at time required by the schedule, two robots should get close enough with each other to observe each other's behavior. This constraint is modeled as a relative distance constraint between two robots at some time instant $j$ (i.e., to require two agents see each other within a certain radius to inspect each other, or to exchange data). The location of the co-observation location should make sure that the ellipsoidal reachability region between each co-observation, considered in later sections \ref{sec:reachability}, has an empty intersection with all forbidden regions.
 We write the function for the constraint as:
\begin{align}
D(\vq) = \overrightarrow{q_{aj}q_{bj}},
\end{align}
where $a,b$ are the indices of the pair of agents required for a mutual inspection. The corresponding constraint set is simply
\begin{equation}
\sZ= \{z \mid z \leq d_{max} \},
\end{equation}
with a projection function:
\begin{equation}\label{eq:co-observation_constraint_update}
\Pi_{\sZ}(z) = \begin{cases}
d_{max}\frac{z}{\norm{z}} &\text{if } \norm{z} > d_{max},\\
z	& otherwise.
\end{cases}
\end{equation}
The locations $q_{aj}$ and $q_{bj}$ where the co-observation performed are computed as part of the optimization.

\subsection{Definition of reachability constraint}\label{sec:reachability}
  In this section, we first provide the definition of an \emph{ellipsoidal reachability region}. We then provide a differentiable map to transform such region in a canonical axis-aligned form where the operator $\Pi_\zeta$ and its differential can be obtained; these operators are then extended to the general case via the aforementioned transform. The overall goal is to define the functions $D(q)$, its differential, and the operator $\Pi_\zeta$ for \emph{ellipsoidal reachability regions} with respect to the forbidden regions that can then be used in the ADMM formulation in section \ref{chapter:ADMM review}.

The reachability region is defined as the set of locations $q(t)$ that a robot can reach between two given fixed positions:

\begin{definition}\label{sec:ellipsoidal definition}
	The \emph{reachability region} for two waypoints $q(t_1)=q_1$, $q(t_2)=q_2$ is defined as the sets of points $q'$ in the workspace such that there exist a trajectory $q(t)$ where $q(t')=q'$, $t_1\leq t' \leq t_2$ and $q(t)$ satisfies the velocity constraint $d(q(t),q(t+1))\leq v_{max}$.
\end{definition}
This region can be analytically bounded via an ellipsoid:
\begin{definition}\label{def:Reachability}
	The \emph{reachability ellipsoid} is defined as the region  $\mathcal{E}(q_1,q_2,t_{1},t_{2})=\{\tilde{q}\in\mathbb{R}^n: d(q_1,\tilde{q})+d(\tilde{q}+q_2)<2a\}$, where $a=\frac{v_{maq}}{2}(t_2-t_1)$.
\end{definition}

@@

The reachability ellipsoid is an over-approximation of the exact reachability region; the difference between the two is due to the discretization of the trajectory, and the fact that $\cE$ does not consider the presence of obstacles.
\subsubsection{Transformation to canonical coordinates}\label{sec:rotation2Standard}
  To simplify the problem, a canonical rigid body transformation is used to transform the ellipse $\cE$ from a global frame $\cF$ to a canonical frame $\cF_\cE$. The latter is defined such that the center of the ellipsoid is located at the origin and the foci are aligned with the first axis of $\cF_\cE$. Since the transformation depends on the two foci, i.e., two waypoints of an agent, the challenge here is to construct the transformation in a differentiable way.
  For the convenience of derivation, we define the coordinate transformation from $\cF$ to $\cF_\cE$ using a rotation $R^\cF_\cE$ and a translation $o^\cF_\cE$, which, to simplify the notation, from now on we simply refer to as $R$ and $o$, respectively. The transformation of a point from the frame $\cF_\cE$ to the frame $\cF$ and its inverse are given by the formulation:
  \begin{equation}
    \Fframe{q}=R\Eframe{q}+o,\label{eq:E to F transform}\quad
    \Eframe{q}=R\transpose(\Fframe{q}-o).
  \end{equation}
  The reverse transformation is $\Eframe{q}=R\transpose( \Fframe{q}-o)$.
